# (PART) Get Started {-}

```{r, include = FALSE}
source("common.R")
```

# R Quickstart

## Get Started with Earth Engine {-}

This Get Started guide is intended as a quick way to start programming with the Earth Engine in R API. For an introductory look at R and more in-depth exercises with the Earth Engine API, see the [tutorials]( ). For recommended R coding style, see the [R Style Guide]( ).

Google Earth Engine allows users to run algorithms on georeferenced imagery and vectors stored on Google's infrastructure. The Google Earth Engine API provides a library of functions which may be applied to data for display and analysis. Earth Engine's [public data catalog](https://developers.google.com/earth-engine/datasets) contains a large amount of publicly available imagery and vector datasets. Private assets can also be created in users' personal folders.

## How to use these docs {-}

The Earth Engine documentation is designed for people familiar with geospatial data analysis. The docs are primarily structured by data type. For example, the left side navigation contains links to sections or pages about important data types such as [Image](), [ImageCollection](), [Feature](), [FeatureCollection](), [Geometry](), [Reducer](), [Chart](), [Join]() and [Array](). This guide only describes only enough about these types to get you started. There are also sections for [machine learning](), specialized or sensor specific algorithms (e.g. [Landsat algorithms]()), [public facing apps]() and important details about [Earth Engine internal workings](). Before diving in to any of that, start here!

## Earth Engine data structures {-}

The two most fundamental geographic data structures in Earth Engine are [Image]() and [Feature]() corresponding to raster and vector data types, respectively. Images are composed of bands and a list of properties. Features are composed of a [Geometry]() and a list of properties. A stack of images (e.g. an image time series) is handled by an [ImageCollection](). A collection of features is handled by a [FeatureCollection](). Other fundamental data structures in Earth Engine include `List`, `Array`, `Date`, `Number` and `String` (learn more about basic data types from [this tutorial](). It is important to remember that these are all server-side objects and are not manipulated the same way as client-side R objects are ([learn more]()).

## Earth Engine algorithms {-}

There are several ways to run operations in the API:

- Calling methods that are attached to objects.
- Calling algorithms.
- Calling Code Editor specific functions.
- Defining new functions.

The *Docs* tab of the Code Editor lists the methods of each API class. For example, the `Image` class has an `add()` method:

```{r, eval=FALSE}
image3 <- image1$add(image2)
```

This method adds the bands of `image2` to the bands of `image1`. The `ee$Algorithms` category contains a list of currently supported algorithms for specialized or domain specific processing. For example, to create topographic layers from an input Digital Elevation Model (DEM):

```{r, eval=FALSE}
terrainImage <- ee$Algorithms$Terrain(dem)
```

Code Editor specific functions include the `Map` and `Export` methods, which control how layers are added to the map panel or exported to Google Drive, respectively. 

The following sections illustrate these concepts for various simple use cases.

*Notation used in the guides:*

Static methods called on an Earth Engine class (for example ee.Image) are written as `Image$staticMethod()`. Methods called on an instance of a class are written as `image$instanceMethod()`. The lowercase image means that a variable named image refers to an instance of the ee.Image class.

## 'Hello World' - R {-}

Printing out information to the console is a basic task for getting information about an object, displaying the numeric result of a computation, displaying object metadata or helping with debugging. The iconic ‘Hello World!’ example in R is:
```{r, eval=FALSE}
print('Hello world!')
```
Copy this line into the RStudio Code Editor and click Run. Note that the output is displayed in the Console tab, on the right or down of the Code Editor. For a more remote sensing relevant example, the following prints the metadata of a Landsat 8 image:

```{r, eval=FALSE}
print(ee$Image('LANDSAT/LC08/C01/T1/LC08_044034_20140318'))
```
Carefully inspect the output in the console to see metadata available for Landsat images.

## Adding data to the map {-}

In addition to printing information to the console, adding data to the Map is the way to visualize geographic data. Use `Map$addLayer()` to do that. In the following example, an Image is instantiated (how to find these images is covered later) using `ee$Image()`, added to the map with `Map$addLayer()` and the map is centered over the image:

```{r, eval=FALSE}
## Load an image.
image <- ee$Image('LANDSAT/LC08/C01/T1/LC08_044034_20140318')

## Center the map on the image.
Map$centerObject(image, 9)

## Display the image.
MapaddLayer(image)
```

The second parameter of `Map$centerObject()` is a zoom level, where higher numbers indicate larger scale (more zoomed in). The parameters for the `Map` functions are described in depth in the API reference accessible from the *Docs* tab. If the appearance of the image is unsatisfactory, configure the display parameters with an additional argument to `Map$addLayer()`. For example:

```{r, eval=FALSE}
## Load the image from the archive.
image <- ee$Image('LANDSAT/LC08/C01/T1/LC08_044034_20140318')

## Define visualization parameters in an object literal.
vizParams <- list(bands = 'B5', 'B4', 'B3',
                  min: 5000, 
                  max: 15000, 
                  gamma: 1.3)

## Center the map on the image and display.
Map$centerObject(image, 9)
Map$addLayer(image, vizParams, 'Landsat 8 false color')
```

Observe that the visualization parameters are defined by an object literal, which includes a list of bands to display, a minimum and maximum digital number and a gamma value. (Learn more about Landsat bands [here](). Learn more about image visualization [here]()).

Use `Map$addLayer()` to add features and feature collections to the map. For example,

```{r, eval=FALSE}
counties <- ee$FeatureCollection('TIGER/2016/Counties')
Map$addLayer(counties, 'counties')
```

## Finding images, image collections and feature collections {-}

Images, image collections, and feature collections are discoverable by searching the Earth Engine Data Catalog. For example, entering ‘Landsat 8’ into the search field results in a list of raster datasets. (The complete listing of Earth Engine datasets is at the [Earth Engine Data Catalog](https://developers.google.com/earth-engine/datasets)). Click on the dataset name to get a brief description, information about the temporal availability, data provider and collection ID.

Alternatively, copy the collection ID and paste it into your code. For example, choose the first result of the ‘Landsat 8’ search and copy the ID as follows:

```{r, eval=FALSE}
collection = ee$ImageCollection('LANDSAT/LC08/C01/T1')
```

Since this is a collection of many images spanning the Earth land surface, finding an individual image in the collection requires filtering in order to narrow down the search. Alternatively, the collection of images can be reduced to a single image using compositing and mosaicking techniques. More on filtering and compositing (see [Reducing]()) in the next sections.

Feature collections are also available through the Data Catalog. Their representation is smaller than image collections, but you will find international, census, watershed, and protected areas boundaries, to name a few. Learn more about importing your own vector datasets [here]().


# Coding Best Practices

To start your journey in mastering R, the following six chapters will help you learn the foundational components of R. I expect that you've already seen many of these pieces before, but you probably have not studied them deeply. To help check your existing knowledge, each chapter starts with a quiz; if you get all the questions right, feel free to skip to the next chapter!

# Debugging

To start your journey in mastering R, the following six chapters will help you learn the foundational components of R. I expect that you've already seen many of these pieces before, but you probably have not studied them deeply. To help check your existing knowledge, each chapter starts with a quiz; if you get all the questions right, feel free to skip to the next chapter!
