# (PART) Specialized Algorithms {-}

```{r, include = FALSE}
source("common.R")
```


# Landsat Algorithms {-}

## Landsat collection structure {-}

The USGS produces data in 3 tiers (categories) for each satellite:

*Tier 1 (T1) - Data that meets geometric and radiometric quality requirements

*Tier 2 (T2) - Data that doesn't meet the Tier 1 requirements

*Real Time (RT) - Data that hasn't yet been evaluated (it takes as much as a month).

See [USGS documentation](https://www.usgs.gov/landsat-missions/landsat-collection-1?qt-science_support_page_related_con=1#qt-science_support_page_related_con) on Collection 1 Tiers for more information.

To allow access to both the validated T1 data and the newest real-time data together, we've grouped the scenes into collections by tier and satellite. Examples for Landsat 8 are as follows:

| ID | Description|
|----|------------|
|LANDSAT/LC08/C01/T1_RT|Landsat 8, Collection 1, Tier 1 + Real Time|
|LANDSAT/LC08/C01/T1|Landsat 8, Collection 1, Tier 1 only|
|LANDSAT/LC08/C01/T2|Landsat 8, Collection 1, Tier 2 only|

Newly acquired scenes are added to the T1_RT collection daily. Once an RT scene gets reprocessed and categorized as either T1 or T2, it will be removed from the T1_RT collection and the new version will be added to the appropriate collection(s). If your work is sensitive to removals or potentially mis-registered scenes, you might want to stick to the T1 collection, but in general, it's very uncommon that any misregistration is large enough to notice on newly acquired scenes.

Each of the above collections contains the raw data (i.e., scaled, at-sensor radiance). In addition, for each collection that contains T1 or T2 images, TOA (top-of-atmosphere reflectance), SR (surface reflectance), and LST (land surface temperature) products are offered. The following table describes the collection ID for TOA and SR/LST collections using Landsat 8 data as an example.

| ID | Description|
|----|------------|
|LANDSAT/LC08/C01/T1_RT_TOA|Landsat 8, Collection 1, Tier 1 + Real Time, TOA|
|LANDSAT/LC08/C01/T1_TOA|Landsat 8, Collection 1, Tier 1 only, TOA|
|LANDSAT/LC08/C02/T1_L2|Landsat 8, Collection 2, Tier 1 only, SR, and LST|
|LANDSAT/LC08/C01/T2_TOA|Landsat 8, Collection 1, Tier 2 only, TOA|
|LANDSAT/LC08/C01/T2_SR|Landsat 8, Collection 1, Tier 2 only, SR (no LST)|

These data exist for Landsat 4, 5, 7, and 8. Replace 'LC08' in the above collection definitions with IDs from the following table to retrieve collections for the various satellites.

| ID | Description|
|----|------------|
|LT04|Landsat 4, Thematic Mapper (TM)|
|LT05|Landsat 5, Thematic Mapper (TM)|
|LE07|Landsat 7, Enhanced Thematic Mapper Plus (ETM+)|
|LC08|Landsat 8, Operational Land Imager (OLI)|

## Landsat collection status {-}

**Pre-Collection 1**: no longer being produced by USGS, unmaintained by Earth Engine, still available in Earth Engine but will eventually be purged.

**Collection 1**: produced by USGS until 2022-01-01, maintained by Earth Engine until 2023-01-01.

**Collection 2**: primary collection produced by USGS, increasing product availability in Earth Engine as ingestion backlog is processed. See the Data Catalog's [Landsat page](https://developers.google.com/earth-engine/datasets/catalog/landsat) for currently available products.

## Landsat processing methods {-}

Earth Engine contains a variety of Landsat specific processing methods. Specifically, there are methods to compute at-sensor radiance, top-of-atmosphere (TOA) reflectance, surface reflectance (SR), cloud score and cloud-free composites.

### At-sensor radiance and TOA reflectance {-}

The ‘raw’ scenes in Earth Engine contain imagery with digital numbers (DNs) that represent scaled radiance. The conversion of DNs to at-sensor radiance is a linear transformation using coefficients stored in scene metadata [(Chander et al. 2009)](https://www.sciencedirect.com/science/article/abs/pii/S0034425709000169). The `ee.Algorithms.Landsat.calibratedRadiance()` method performs this conversion. Conversion to TOA (or at-sensor) reflectance is a linear transformation that accounts for solar elevation and seasonally variable Earth-Sun distance. The TOA conversion is handled by the `ee.Algorithms.Landsat.TOA()` method. The TOA method converts thermal bands to brightness temperature. See [Chander et al. (2009)](https://www.sciencedirect.com/science/article/abs/pii/S0034425709000169) (or [this USGS site](https://landsat.usgs.gov/using-usgs-landsat-8-product) for Landsat 8) for more information about computing TOA reflectance or brightness temperature. The following example shows conversion from raw data to radiance and TOA reflectance for a Landsat 8 image:

```{r, eval=FALSE}
# Load a raw Landsat scene and display it.
raw = ee$Image('LANDSAT/LC08/C01/T1/LC08_044034_20140318')
Map$centerObject(raw, 10)
Map$addLayer(raw, {bands = ['B4', 'B3', 'B2'], min = 6000, max = 12000}, 'raw')

# Convert the raw data to radiance.
radiance = ee$Algorithms$Landsat$calibratedRadiance(raw)
Map$addLayer(radiance, {bands = ['B4', 'B3', 'B2'], max = 90}, 'radiance')

# Convert the raw data to top-of-atmosphere reflectance.
toa = ee$Algorithms$Landsat$TOA(raw)

Map$addLayer(toa, {bands = ['B4', 'B3', 'B2'], max: 0.2}, 'toa reflectance')
```

### Surface reflectance {-}

Landsat surface reflectance (SR) data are available in Earth Engine as a copy of the USGS Collection 2, Level 2 archive. Note that Landsat 4, 5, and 7 SR data are generated using the LEDAPS algorithm, while Landsat 8 SR data are generated using the LaSRC algorithm. [Learn about these algorithms and their differences from USGS.](https://www.usgs.gov/landsat-missions/landsat-collection-2-surface-reflectance)

You can access a USGS Collection 2, Level 2 Landsat 8 image like this:

```{r, eval=FALSE}
srImage = ee$Image('LANDSAT/LC08/C02/T1_L2/LC08_044034_20201028')
```

The surface reflectance datasets for Collection 2 Landsat 4 through 8 are:

```{r, eval=FALSE}
surfaceReflectanceL4 = ee$ImageCollection('LANDSAT/LT04/C02/T1_L2')
surfaceReflectanceL5 = ee$ImageCollection('LANDSAT/LT05/C02/T1_L2')
surfaceReflectanceL7 = ee$ImageCollection('LANDSAT/LE07/C02/T1_L2')
surfaceReflectanceL8 = ee$ImageCollection('LANDSAT/LC08/C02/T1_L2')
```

### Simple cloud score {-}

For scoring Landsat pixels by their relative cloudiness, Earth Engine provides a rudimentary cloud scoring algorithm in the `ee$Algorithms$Landsat$simpleCloudScore()` method. The following example uses the cloud scoring algorithm to mask clouds in a Landsat 8 image:

```{r, eval=FALSE}
# Load a cloudy Landsat scene and display it.
cloudy_scene = ee$Image('LANDSAT/LC08/C01/T1_TOA/LC08_044034_20140926')
Map$centerObject(cloudy_scene)
Map$addLayer(cloudy_scene, {bands = ['B4', 'B3', 'B2'], max: 0.4}, 'TOA', false)

# Add a cloud score band.  It is automatically called 'cloud'.
scored = ee$Algorithms$Landsat$simpleCloudScore(cloudy_scene)

# Create a mask from the cloud score and combine it with the image mask.
mask = scored$select(['cloud'])$lte(20)$

# Apply the mask to the image and display the result.
masked = cloudy_scene$updateMask(mask)
Map$addLayer(masked, {bands = ['B4', 'B3', 'B2'], max: 0.4}, 'masked')

```

If you run this example in the Code Editor, try toggling the visibility of the TOA layers to compare the difference between the masked and unmasked imagery. (See this section of the Code Editor docs for instructions on how to do that). Observe that the input to `simpleCloudScore()` is a single Landsat TOA scene. Also note that `simpleCloudScore()` adds a band called ‘cloud’ to the input image. The cloud band contains the `cloud` score from 0 (not cloudy) to 100 (most cloudy). The previous example uses an arbitrary threshold (20) on the cloud score to mask cloudy pixels. To apply this algorithm to an Earth Engine mosaic of Landsat scenes, set the `SENSOR_ID` property:

```{r, eval=FALSE}
# Load a Landsat 8 composite and set the SENSOR_ID property.
mosaic = ee$Image(ee$ImageCollection('LANDSAT/LC8_L1T_8DAY_TOA')$first())
  .set('SENSOR_ID', 'OLI_TIRS')

# Cloud score the mosaic and display the result.
scored_mosaic = ee$Algorithms$Landsat.simpleCloudScore(mosaic)
Map$addLayer(scored_mosaic, {bands = ['B4', 'B3', 'B2'], max: 0.4},
    'TOA mosaic', false)
```

`SENSOR_ID` is a property of individual images. When Earth Engine makes a mosaic of many images, it has to throw out individual image metadata, including the `SENSOR_ID` property. To cloud score a mosaic, Earth Engine looks for that property and can't find it, resulting in an error. Set the property manually to avoid that. The sensor IDs of Landsat 5, 7 and 8 are 'TM', 'ETM+' and 'OLI_TIRS', respectively.

### Simple composite {-}

For creating simple cloud-free Landsat composites, Earth Engine provides the `ee.Algorithms.Landsat.simpleComposite()` method. This method selects a subset of scenes at each location, converts to TOA reflectance, applies the simple cloud score and takes the median of the least cloudy pixels. This example creates a simple composite using default parameters and compares it to a composite using custom parameters for the cloud score threshold and the percentile:

```{r, eval=FALSE}
# Load a raw Landsat 5 ImageCollection for a single year.
collection = ee$ImageCollection('LANDSAT/LT05/C01/T1')
    .filterDate('2010-01-01', '2010-12-31')

# Create a cloud-free composite with default parameters.
composite = ee$Algorithms$Landsat$simpleComposite(collection)

# Create a cloud-free composite with custom parameters for
# cloud score threshold and percentile.
customComposite = ee$Algorithms$Landsat$simpleComposite({
  collection = collection,
  percentile = 75,
  cloudScoreRange = 5
})

# Display the composites.
Map$setCenter(-122.3578, 37.7726, 10)
Map$addLayer(composite, {bands = ['B4', 'B3', 'B2'], max: 128}, 'TOA composite')
Map$addLayer(customComposite, {bands = ['B4', 'B3', 'B2'], max: 128},
    'Custom TOA composite')
```

Note that the input to the simple composite is a collection of raw imagery. Also note that by default, reflective band output is reflectance scaled to 8-bits and thermal band output is Kelvin minus 100, to fit in the 8-bit range. You can change this behavior by setting the `asFloat` parameter to true, to get un-scaled, un-shifted float output.

# Sentinel-1 Algorithms

To start your journey in mastering R, the following six chapters will help you learn the foundational components of R. I expect that you've already seen many of these pieces before, but you probably have not studied them deeply. To help check your existing knowledge, each chapter starts with a quiz; if you get all the questions right, feel free to skip to the next chapter!

# Resampling and Reducing Resolution  

To start your journey in mastering R, the following six chapters will help you learn the foundational components of R. I expect that you've already seen many of these pieces before, but you probably have not studied them deeply. To help check your existing knowledge, each chapter starts with a quiz; if you get all the questions right, feel free to skip to the next chapter!
